### 2020-03-18 ChapterBackground/chapterBackground.tex
Fixed: Duplicate reference
Fixed: Proof-read of background

\subsubsection{Peer-to-Peer}
Peer-to-Peer, often referred to as`P2P', is a type of architecture in which the game is executed and game state is managed entirely by the peers, although peers are often only responsible for a subset of the game state, such as the entities they are currently interacting with. A peer-to-peer architecture is shown in Fig. \ref{fig_P2P}. Peers are all responsible for message forwarding to each other. In other words, peers act as both server and client. Peer-to-Peer reduces the cost for the game provider of maintaining expensive servers and can reduce latency between players, as communication is performed directly between peers, rather than communicating through a server. In addition, it provides scalability in terms of compute power, as the more peers that join, the more resources available to the system. However, as more peers join, the greater the communication overhead is between peers. In addition, peer-to-peer is prone to the problems of fault-tolerance, data consistency and cheating by players. A peer is more likely to fail or experience unscheduled disconnection than a dedicated server. As each peer is responsible for the game state, it is possible for communication errors to occur, resulting in two peers with different game states or a peer maliciously altering the game state or the messages being sent to other peers in order to cheat. The fault-tolerance requirements of peer-to-peer can be reduced in short-session-based games, making these architectures better suited to those games \cite{P2PForMMOs}. Peer-to-Peer architectures have received much research attention in previous years, with many solutions that attempt to address the problems above, through the use of techniques discussed in \cite{P2PForMMOs}.

@@

\subsubsection{Hybrid Peer-to-Peer}
Hybrid Peer-to-Peer architectures are a combination of client-server and peer-to-peer architectures. The peer-to-peer part of the system can be responsible for different aspects of the system, including Cooperative Message Dissemination, State Distribution, Basic Server Control and distributing software updates. \cite{P2PForMMOs}. Hybrid Peer-to-Peer enable trade-offs between consistency control and scalability are enabled, depending on how the system is divided between server and peer-to-peer.
%These can be broadly classified into two types: Homogeneous, in which all nodes have similar responsibilities and Heterogeneous, in which some peers be

@@

In the `shards' method, complete and separate instances of the game world exist, each maintained by one serve. Each server is responsible for a different set of clients. The game can be scaled through the introduction of more servers, supporting more clients. However, as each instance of the game world is separate, there can be no interaction in the game world between players on different shards \cite{P2PForMMOs}.

In the game world division method, there is a single instance of the game world and the workload is distributed either through each server being responsible for a region or a set of players. Division methods are discussed further in \ref{distrubted_virtual_env}. 
a single server can be overloaded if the workload is too high, for example, by a large number of players in a region being managed by a single server. Dynamic load-balancing schemes can be employed to improve distribution between servers. The disadvantage of this approach is the need for complex hand-offs when players migrate or interact between servers and can lead to consistency problems between servers as latency will always exist between servers \cite{P2PForMMOs}.

@@

The rates chosen for network tick rates depend on a variety of factors, including the required level of accuracy and bandwidth restrictions. Higher tick rates enable higher accuracy, but at the cost of more bandwidth being required. Different genres of games have different accuracy demands. For example, in first-person-shooters (FPS), higher tick rates result in a higher mean shooting accuracy for players \cite{lee2015evaluation}. Whereas, genres like real-time-strategy (RTS) games have more relaxed accuracy requirements. In the RTS Age of Empires, latency only became noticeable if it exceeded $500ms$, meaning network tick rates can be very low without being noticeable to the player \cite{pisan2004challenges}. Tick rates of popular network games range from 128Hz for CS:GO to 20Hz for Minecraft \cite{metzger2016comprehensive}.

@@

\subsection{Protocols}
Games use both TCP (Transmission Control Protocol) and UDP (User Datagram Protocol). Most games use the UDP network protocol for transmission of short and time sensitive data and TCP for long and content-aware data, providing services such as matchmaking \cite{2018NetworkTraffic, 2012PacketLevelTraffic}. UDP offers best effort communication . Unlike TCP, UDP is a connection-less protocol and does not guarantee data delivery, but suffers from less latency. In many situations in online games, lost packets are not critical, for example in an online game in which  the server informs each client of the positions of each player, there are a large number of updates, meaning a missed update is not critical \cite{pisan2004challenges}. 
%Network play in online games has to be constructed around latency and packet-loss

\section{RakNet}
RakNet is the gaming network library used in the development of the implementation of this project. Features of RakNet used by this study's implementation will now be described.

@@

The former is a very common issue in online gaming, particularly due to high latency and limited bandwidth between servers and client, necessitating the need for client-side correction techniques. The most common technique used is dead-reckoning, shown in Fig. \ref{fig_DeadReckoning}. Dead-reckoning's use in games is well researched, including the suitability for different genres \cite{2002DeadReckoningSuitability}, the accuracy of dead-reckoning \cite{2004AccuracyDeadReckoning}, its effects on fairness \cite{2005FairnessDeadReckoning} and enhancements of the technique, such as \cite{2012EnhancedDeadReckoning} which takes into account environmental objects when dead-reckoning is performed.

@@

\begin{enumerate}
	\item Server-authoritative physics - the server is responsible for simulating the physics aspect of the game
	\item Client-authoritative physics - the physics is simulated on the clients in a distributed manner
\end{enumerate}

@@

Server-authoritative physics multi-server systems creates complex problems at the boundaries between servers. These boundary problems include:
\begin{itemize}
\item The handing off (migration) of physical entities between servers
\item Maintaining consistent states of physical entities between servers
\item The interaction of entities between servers, i.e. objects being simulated on different servers that should physically interact with each other
\end{itemize}

@@

This study aims to address the problems of consistency in distributing real-time physics between servers in multi-server networked games. This would allow for physics to play a larger role in networked games, including having significant influences on gameplay.

@@

In the case of architectures not using shards, Interest Management is required to prevent message passing growing polynomially as players increase (e.g. \cite{Bezerra2008} and \cite{LoadBalancingforMMOs}).
\cite{Bezerra2008} proposes the A3 algorithm, an interest management technique for distributed simulations aiming to significantly reduce the necessary bandwidth required between servers. A3 uses a combination of a circular area of interest and field of view combined with a relevance gradient. \cite{LoadBalancingforMMOs} proposes a Behavioural Interest Management Technique that allocates resources based on player interactions. Auras (an area of interest/influence) are used to determine player message exchanges, reducing message passing while promoting player number scalability.

@@

Interest Management is relevant to our work as it enables the minimisation of network overhead in distributed virtual environments. It is important to note that objects within a virtual space are only affected by other objects in their region and neighbouring regions or objects within their area of influence. Aura Projection uses the concept of auras to minimise message passing between servers while still allowing time-space consistency to be maintained.

In addition to reducing the network traffic between servers, interest management can also be utilised to reduce the required network traffic between client and servers. In terms of physics being simulated on server(s), clients will be most concerned with objects within close proximity to them and less concerned with objects beyond the range in which can be physically interacted with. In addition, clients will be more concerned with objects in their field of view than objects that can't currently be seen. An example of an interest management technique using distance and field of view is \cite{Bezerra2008}. \cite{Bezerra2008} proposes the A3 algorithm, an interest management technique for distributed simulations aiming to significantly reduce the necessary bandwidth required in a distributed game server. A3 uses a combination of a circular area of interest and field of view combined with a relevance gradient.

Through the use of client interest management it would be possible to simulate a persistent world far larger than could be simulated or rendered on a single machine, yet allow a player unrestricted, seamless access to the entire world.

@@

Examples of Distributed Real-Time Physics that use multiple instances of a physics engine include SpatialOS \cite{SpatialOS} and Aether Engine \cite{AetherEngine}. SpatialOS spatially partitions the world into regions, each region is a separate instance of a physics engine. Where regions meet, there is an overlapping area, allowing objects hosted on different servers to interact. However, the specifics of the techniques used by SpatialOS and Aether Engine are not described in any literature. The demonstration video of SpatialOS exhibits unnatural object "jitter", which is possibly a result of network latency.

@@

Cloud deployment models are categorised into the following:
\begin{itemize}
\item Private cloud - Set up within an organisation's internal data centre. This is much more secure than the public cloud due to its internal only exposure and access is limited to designated users within an organisation \cite{2012CloudComputing}.
\item Public cloud - Resources are provided by a third-party provider. Users request resources in a self-service manner over the Internet. Users are typically charged on a pay-per-use model. Public clouds are less secure than private cloud as users have more responsibility in ensuring the security of their applications and data from potential malicious attacks from other users in the cloud. Public clouds offer far greater resources than a private cloud, allowing spikes in demand to be provided for  \cite{2012CloudComputing}.
\item Hybrid cloud - A private cloud linked to one or more public clouds. This provides solutions with more secure control over data and applications while still allowing public access over the internet \cite{2012CloudComputing}.
\end{itemize}

@@

This study is concerned with inter-region and inter-AZ latencies as these are the latencies that affect server-server communication and the correctness of the system. Lower latency is ideal for the deployment of a real-time distributed physics system, in the case of AP, lower latency means smaller auras leading to better partitioning and subsequent performance. However, inter-AZ deployment may not always be possible or desired (e.g. lack of availability of resources in one AZ or to reduce latency between users and some of the servers in the system in the case of geographically diverse users). Therefore the latency experiments used Inter-AZ latency and low Inter-region latency values.  The lowest latency observed was from the inter-AZ category, i.e. servers hosted in the same geographical region at $<2ms$. Inter-region latency measured in the range of ($7-302ms$).

@@

The chosen values for the packet-loss correctness experiment range from $0\%-20\%$. These values cover the entire range of expected packet-loss values while running in the cloud, as well as beyond those values, in order to determine what impact severe packet-loss would have on the correctness of AP and establish the trend increasing packet-loss has.

@@

Popular existing commercial game engines, such as Unity \cite{Unity} and Unreal Engine \cite{Unreal} , are based on the Monolith architecture. However, for game engines to take greater advantage of cloud deployment a move to a microservice architecture will be needed. This would allow for independent horizontal scaling and resource provision management of the different systems within a game engine, such as graphics, physics and AI. Despite the recent commercial interest in microservice architecture, there has been no academic research into the use of microservices for game engines. The only notable exception to this is \cite{vaha2017applying}, however, this has a limited focus to the application of microservices to the genre of Massively Multiplayer Online Role-Playing Games (MMORPG).

A microservice architecture for a game engine would create a more cost efficient approach to cloud-hosted games and simulations, while also creating the potential for the scalability of individual systems on an on-demand basis. This study's problem is concentrated on separating out the physics system from the game engine and distributing it across multiple servers while maintaining the real-time streamed service to players. Through this distribution, this study aims to demonstrate that the physics aspect of a game engine can be horizontally scaled through the use of cloud computing.

@@

In recent years, streamed gaming has become an active area of research, with many studies focusing on the effects of latency on the quality of service, e.g. \cite{CloudGamingArchPerf, chen2011measuring, jarschel2011evaluation, lee2012all, choy2012brewing, chen2013quality}. The challenges in streamed gaming that remain are both technical as well as economical \cite{2016SurveyOnCloudGaming}.

>>>>

\subsubsection{Peer-to-Peer}
Peer-to-Peer, often referred to as`P2P', is a type of architecture in which the game is executed and game state is managed entirely by the peers, although peers are often only responsible for a subset of the game state, such as the entities they are currently interacting with. A peer-to-peer architecture is shown in Fig. \ref{fig_P2P}. Peers are all responsible for message forwarding to each other. In other words, peers act as both server and client. Peer-to-Peer reduces the cost for the game provider of maintaining expensive servers and can reduce latency between players, as communication is performed directly between peers, rather than communicating through a server. In addition, it provides scalability in terms of compute power, as the more peers that join, the more resources available to the system. However, as more peers join, the greater the communication overhead is between peers. In addition, peer-to-peer is prone to the problems of fault-tolerance, data consistency and cheating by players. A peer is more likely to fail or experience unscheduled disconnection than a dedicated server. As each peer is responsible for the game state, it is possible for communication errors to occur, resulting in two peers with different game states or a peer maliciously altering the game state or the messages being sent to other peers in order to cheat. The fault-tolerance requirements of short-session-based games are lower than persistent games, making peer-to-peer architectures better suited to short-session-based games \cite{P2PForMMOs}. Peer-to-Peer architectures have received much research attention in previous years, with many solutions that attempt to address the problems above, through the use of techniques discussed in \cite{P2PForMMOs}.

@@

\subsubsection{Hybrid Peer-to-Peer}
Hybrid Peer-to-Peer architectures are a combination of client-server and peer-to-peer architectures. The peer-to-peer part of the system can be responsible for different aspects of the system, including Cooperative Message Dissemination, State Distribution, Basic Server Control and distributing software updates. \cite{P2PForMMOs}. Hybrid Peer-to-Peer enables trade-offs between consistency control and scalability, determined by how the system is divided between server and peer-to-peer. The more responsibility peers have, the greater the scalability but at a loss of consistency control.
%These can be broadly classified into two types: Homogeneous, in which all nodes have similar responsibilities and Heterogeneous, in which some peers be

@@

In the `shards' method, complete and separate instances of the game world exist, each maintained by one server. Each server is responsible for a different set of clients. The game can be scaled through the introduction of more servers, supporting more clients. However, as each instance of the game world is separate, there can be no interaction in the game world between players on different shards \cite{P2PForMMOs}.

In the game world division method, there is a single instance of the game world and the workload is distributed either through each server being responsible for a region or a set of players. Division methods are discussed further in \ref{distrubted_virtual_env}. Even with the workload distributed between servers, one server may be become overloaded if the workload is too high, for example, by a large number of players in a region being managed by a single server. Dynamic load-balancing schemes can be employed to improve distribution between servers and prevent servers being overloaded when resources are available on other servers. The disadvantage of distributed servers is the need for complex hand-offs when players migrate or interact between servers and can lead to consistency problems between servers as latency will always exist between servers \cite{P2PForMMOs}.

@@

The rates chosen for network tick rates depend on a variety of factors, including the required level of accuracy and bandwidth restrictions. Higher tick rates enable higher accuracy, but at the cost of more bandwidth being required. Different genres of games have different accuracy demands. For example, in first-person-shooters (FPS), higher tick rates result in a higher mean shooting accuracy for players \cite{lee2015evaluation}. Whereas, genres like real-time-strategy (RTS) games have more relaxed accuracy requirements. In the RTS Age of Empires, latency only becomes noticeable if it exceeds $500ms$, meaning network tick rates can be very low without being noticeable to the player \cite{pisan2004challenges}. Tick rates of popular network games range from 128Hz for CS:GO to 20Hz for Minecraft \cite{metzger2016comprehensive}.

@@

\subsection{Protocols}
Games use both TCP (Transmission Control Protocol) and UDP (User Datagram Protocol). Most games use the UDP network protocol for transmission of short and time-sensitive data and TCP for long and content-aware data, providing services such as matchmaking \cite{2018NetworkTraffic, 2012PacketLevelTraffic}. UDP offers best effort communication. Unlike TCP, UDP is a connection-less protocol and does not guarantee data delivery, but suffers from less latency. In many situations in online games, lost packets are not critical, for example in an online game in which  the server informs each client of the positions of each player, there are a large number of updates, meaning a missed update is not critical \cite{pisan2004challenges}. 
%Network play in online games has to be constructed around latency and packet-loss

\section{RakNet}
RakNet is the gaming network library used in the development of the implementation of AP. Features of RakNet used by AP will now be described.

@@

The former is a very common issue in online gaming, particularly due to high latency and limited bandwidth between servers and client, necessitating the need for client-side correction techniques. The most common technique used is dead-reckoning, shown in Fig. \ref{fig_DeadReckoning}. The use of dead-reckoning in games is a well researched topic, including the suitability for different genres \cite{2002DeadReckoningSuitability}, the accuracy of dead-reckoning \cite{2004AccuracyDeadReckoning}, its effects on fairness \cite{2005FairnessDeadReckoning} and enhancements of the technique, such as \cite{2012EnhancedDeadReckoning} which takes into account environmental (static) objects when dead-reckoning is performed.

@@

\begin{enumerate}
	\item Server-authoritative physics - the server is responsible for simulating the physics aspect of the game.
	\item Client-authoritative physics - the physics is simulated on the clients in a distributed manner.
\end{enumerate}

@@

Server-authoritative physics in distributed server systems creates complex problems at the boundaries between servers. These boundary problems include:
\begin{itemize}
\item The handing off (migration) of physical entities between servers, e.g. when an object traverses between server regions.
\item Maintaining consistent states of physical entities between servers, i.e. servers will need to be aware of the state of objects being simulated on a different server and this state needs to be kept consistent between the two servers. 
\item The interaction of entities between servers, i.e. objects being simulated on different servers that should physically interact with each other.
\end{itemize}

@@

This study aims to address the problems of consistency in distributing real-time physics between servers in multi-server networked games. Solving this problem would allow for physics to play a larger role in networked games, including having significant influences on gameplay.

@@

In the case of architectures not using shards, Interest Management is required to prevent message passing growing polynomially as players increase (e.g. \cite{Bezerra2008} and \cite{LoadBalancingforMMOs}).

@@

Interest management is a well researched topic and enhancements to it exist, such as \cite{Bezerra2008} which proposes the A3 algorithm. A3 uses a combination of a circular area of interest and field of view combined with a relevance gradient. \cite{LoadBalancingforMMOs} proposes a Behavioural Interest Management Technique that allocates resources based on player interactions. Auras (an area of interest/influence) are used to determine player message exchanges, reducing message passing while promoting player number scalability.

In addition to reducing the network traffic between servers, interest management can also be utilised to reduce the required network traffic between client and servers. In terms of physics being simulated on server(s), clients will be most concerned with objects within close proximity to them and less concerned with objects beyond the range in which can be physically interacted with. In addition, clients will be more concerned with objects in their field of view than objects that can't currently be seen. An example of an interest management technique using distance and field of view is the A3 algorithm \cite{Bezerra2008}. Through the use of client interest management it would be possible to simulate a persistent world far larger than could be simulated or rendered on a single machine, yet allow a player unrestricted, seamless access to the entire world.

Interest Management is relevant to this study as it enables the minimisation of network overhead in distributed virtual environments. It is important to note that objects within a virtual space are only affected by other objects in their region and neighbouring regions or objects within their area of influence. AP uses the concept of auras to minimise message passing between servers while still allowing time-space consistency to be maintained.

@@

Examples of Distributed Real-Time Physics that use multiple instances of a physics engine include SpatialOS \cite{SpatialOS} and Aether Engine \cite{AetherEngine}. SpatialOS spatially partitions the world into regions, each region is a separate instance of a physics engine. Where regions meet, there is an overlapping area, allowing objects hosted on different servers to interact. However, the specifics of the techniques used by SpatialOS and Aether Engine are not described in any literature. The demonstration video of SpatialOS exhibits unnatural object "jitter", which is possibly a result of network latency and collision inconsistencies.

@@

Cloud deployment models are categorised into the following:
\begin{itemize}
\item Private cloud - Set up within an organisation's internal data centre. This is much more secure than the public cloud due to its internal only exposure and access is limited to designated users within an organisation \cite{2012CloudComputing}.
\item Public cloud - Resources are provided by a third-party provider. Users request resources in a self-service manner over the Internet. Users are typically charged on a pay-per-use model. Public cloud services are less secure than private cloud as users have more responsibility in ensuring the security of their applications and data from potential malicious attacks from other users in the cloud. Public cloud services offer far greater resources than a private cloud, allowing spikes in demand to be provided for  \cite{2012CloudComputing}.
\item Hybrid cloud - A private cloud linked to one or more public clouds. This provides solutions with more secure control over data and applications while still allowing public access over the internet \cite{2012CloudComputing}.
\end{itemize}

@@

This study is concerned with inter-region and inter-AZ latencies as these are the latencies that affect server-server communication and the correctness of the system. Lower latency is ideal for the deployment of a real-time distributed physics system, in the case of AP, lower latency means smaller auras leading to better partitioning and subsequent performance. However, inter-AZ deployment may not always be possible or desired (e.g. lack of availability of resources in one AZ or to reduce latency between users and some of the servers in the system in the case of geographically diverse users). Therefore the latency experiments used inter-AZ latency and low inter-region latency values. The lowest latency observed was from the inter-AZ category, i.e. servers hosted in the same geographical region at $<2ms$. Inter-region latency measured in the range of ($7-302ms$).

@@

The chosen values for the packet-loss correctness experiment range from $0\%-20\%$. These values cover the entire range of expected packet-loss values while running in the cloud, as well as beyond those values, in order to determine what impact severe packet-loss would have on the correctness of AP and establish any trend increasing packet-loss has.

@@

Popular existing commercial game engines, such as Unity \cite{Unity} and Unreal Engine \cite{Unreal}, are based on the Monolith architecture. However, for game engines to take greater advantage of cloud deployment a move to a microservice architecture will be needed. This would allow for independent horizontal scaling and resource provision management of the different systems within a game engine, such as graphics, physics and AI. Despite the recent commercial interest in microservice architecture, there has been no academic research into the use of microservices for game engines. The only notable exception to this is \cite{vaha2017applying}, however, this has a limited focus to the application of microservices to the genre of Massively Multiplayer Online Role-Playing Games (MMORPG).

A microservice architecture for a game engine would create a more cost-efficient approach to cloud-hosted games and simulations, while also creating the potential for the scalability of individual systems on an on-demand basis. This study's problem is concentrated on separating out the physics system from the game engine and distributing it across multiple servers while maintaining the real-time streamed service to players. Through this distribution, this study aims to demonstrate that the physics aspect of a game engine can be horizontally scaled through the use of cloud computing.

@@

In recent years, streamed gaming has become an active area of research, with many studies focusing on the effects of latency on the quality of service, e.g. \cite{CloudGamingArchPerf, chen2011measuring, jarschel2011evaluation, lee2012all, StormInCloudGaming, chen2013quality}. The challenges in streamed gaming that remain are both technical as well as economical \cite{2016SurveyOnCloudGaming}.
